The Array of Long Baseline Antennas for Taking Radio Observations from the Sub-Antarctic

### 2020-10-08 albatros_paper.tex
HCC removed old comments from main latex file, removed unused references from bib file
added acknowledgement
started responses to referee

% do we want/need to say anything about the omnidirectional primary beam?

@@

% EE citing the noise figure of the GALI-74+ is particulary misleading because of the large mismatch loss, which directly scales up the noise figure in the HF band.

%% It seems to me that this is too much detail to go into in describing an already published design.
% All the front end components were incorporated for in a double sided
% printed circuit board (PCB) as shown in \autoref{Fig:Balun} and the
% block diagram is shown in \autoref{Fig:Balun Schematic}. The
% Monolithic Microwave Integrated Circuits (MMICs) is the used design
% for the PCB. One side of the PCB is populated with components and
% the other side is a solid copper ground plane aperiodically stitched
% to the grounded copper on the side populated with components. The
% receiver system is made up of the active balun, filter and the gain
% stage that connects to the \SI{100}{\metre} coaxial cable which is
% connected to the back end \cite{2012PASP..124.1090H}.

% The input impedance (Z\textsubscript{o}) of \SI{50}{\ohm} is
% introduced to the dipole by the active balun. The signal is then fed
% through an amplifier which amplifies it by \SI{+24}{\decibel} of
% gain. The balanced signal is then converted to unbalanced through a
% 180 \degree hybrid coupler. The band pass filter (BPF) receives the
% single ended signal in order for it to reject all the frequencies
% which are not within the range of interest. The signal gets fed to a
% second amplifier which again amplies it by \SI{+24}{\decibel} of gain
% and the output impedance of the FEE is matched to a \SI{50}{\ohm}
% coaxial cable. The bias tee is responsible for providing power to the
% FEE and extracts the RF signal by the use of the coaxial cable. This
% unit has an overall gain of $\approx$ \SI{35}{\decibel} and an overall
% noise figure of $\approx$ \SI{2.7}{\decibel} to $\approx$
% \SI{2.9}{\decibel} \cite{Memo35}.

@@

% data rate ~ 17MB/hour, with compressed scio files

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5

@@

% HCC: we probably don't need this figure
% \begin{figure}[h]
% 	\begin{center}
% 		\includegraphics[width=0.5\linewidth]{Figures/balun.jpg}
% 		\caption{Unenclosed FEE mounted on the ALBATROS-EGG antenna supporting structure.} 
% 		\label{Fig:Balun}
% 	\end{center}
% \end{figure}

@@

The back-end readout electronics are housed in a Faraday cage located
50~m away from the antennas. The analog signal chain consists of a
pair of high- and low-pass filters (Mini-Circuits ZFHP-1R2+ and
SLP-150+) that together band-limit the signal to 1.2--140~MHz, and the
filters are followed by a Mini-Circuits ZFL-500+ amplifier.  In
contrast to the two-element pathfinder, the low-pass cutoff is
increased from 81~MHz to 140~MHz to capture downlink signals at
137--138~MHz from the ORBCOMM satellite constellation.  For the final
\albatros\ array, the ORBCOMM signals provide a convenient means for
synchronization across the antenna stations, serving as a backup to
the GPS timing discussed below.  Preliminary lab tests suggest that on
time scales of $\sim30$~s, relative timing between different antenna
stations can be measured to an accuracy of better than a few tenths of
a nanosecond using a single satellite.  The corresponding phase error
at 10~MHz is $\lesssim1^\circ$.  With publicly available orbits and
multiple satellites typically within the field of view, we expect that
ORBCOMM baseband data saved at the same time as the astronomical data
can provide offline synchronization of the \albatros\ stations.

As with the two-element pathfinder, a SNAP board digitizes each of the
two RF signals at 250~Msamp/s. In the autonomous station
configuration, the SNAP board ADCs are locked to a 10-MHz reference
produced by a Trimble Thunderbolt~E GPS-disciplined clock module.  The
SNAP board FPGA computes two data products: 1) channelized baseband
data for each polarization over tunable frequency windows within the
0--125~MHz operating range, with the options of 1-, 2-, or 4-bit
compression, and 2) auto- and cross-spectra from the two polarizations
over the full 0--125~MHz span, accumulated over few-second intervals.
An RPi~3B+ controls the SNAP board and and receives the auto- and
cross-spectra via GPIO connections, and the spectra are saved to an
on-board SD card.  The baseband data are passed from the SNAP board to
the RPi via ethernet and written to external hard drives.  The
introduction of gigabit ethernet with the RPi~3B+ model has enabled
the high data throughput associated with writing baseband.  As a
benchmark, 1-bit baseband recording of two polarizations over 10~MHz
of bandwidth yields an approximate data rate of 5~MB/s, or 0.4~TB/day.

@@

% To get the expected SNR from a 1-bit correlator, multiply the SNR
% derived from the radiometer equation by $\epsilon_{corr}$.

@@

% One might worry that at high signal levels, the saturation of the
% 1-bit cross correlation would lead to the noise exploding.  In
% practice the rolloff in sensitivity is relatively mild, especially in
% the case of long baselines.  It can be expressed in terms of the
% correlator efficiency $\epsilon_{corr}$, which is defined to be the
% ratio of the measured signal-to-noise ratio to the ideal (infinite
% precision) SNR.  As seen in Figure \ref{fig:1bit_efficiency},
% $\epsilon_{corr}$ only drops by a factor of $\sim$2 as the signal
% power goes from 0 to a few times the noise power.  We stress that the
% signal level in question is the part that correlates between the two
% antennas.  In the case of long baselines (where long means the fringe
% spacing is small compared to the primary beam/antenna response
% pattern) that are not dominated by a single bright source, the signal
% power will always be smaller than the noise power, and so
% $\epsilon_{corr}$ is unlikely to drop below $\sim$0.5\footnote{A quick
%   estimate can be made by analogy to the behavior of dishes in the UV
%   plane in the flat sky approximation.  While all power in the UV
%   plane (plus any system noise) goes into an individual antenna's
%   electric field, the correlated part only has contribution from the
%   area in the UV plane within the width of the UV-space primary beam
%   of the UV-space coordinate.  If the sky can be described as a
%   Gaussian random field, in the most pessimistic case this is roughly
%   the aperture filling factor of the antenna pair.  In our case, where
%   the antennas are approximately dipolar, the filling factor will be
%   the square of the wavelength over the baseline length.  As long as
%   the baselines are several wavelengths or longer, the correlated
%   power will be small, and the correlator efficiency will be close to
%   the ideal 1-bit value of $\frac{2}{\pi}$.}.

@@

% note to self: look up time of day that corresponds to LST period
% where fringes appear to extend below 10 MHz, see if that's night
% time

@@

% In the overplotted mean spectra, the baseband data are shown with
% finer, 14.6-kHz resolution to illustrate the additional information
% that is not available in the accumulated spectra.  For these mean
% spectrum plots, the baseband data are processed by
% inverting\footnote{\url{https://github.com/jrs65/pfb-inverse}} the
% polyphase filter bank to recover time-domain data, and the
% forward-direction polyphase filter bank is then applied to
% concatenated timestreams of an appropriate total length to set the
% desired frequency resolution.

@@

We gratefully acknowledge the National Research Foundation (grant
number 110929), the South African National Antarctic Programme, the
Natural Sciences and Engineering Research Council of Canada (grant
number RGPNS-2019-534549), and the Polar Continental Shelf Program for
providing funding and logistical support for our research program.
The financial assistance of the South African SKA Project (SKA SA)
towards this research is hereby acknowledged (\url{www.ska.ac.za}).
We additionally extend our sincere gratitude to the staff at SKA SA
for hosting our lab tests.  We thank the South African National Space
Agency for their technical support, the crew of Ultimate Heli for
safely delivering us and our cargo, and all of the Marion takeover and
winter team members who provided invaluable advice and field help.
The authors also wish to thank members of the LWA and RAPID
collaborations for useful discussions, and Raul Monsalve for helping
make our first MARS campaign a success.

>>>>

@@

@@

@@

@@

@@

The back-end readout electronics are housed in a Faraday cage located
50~m away from the antennas. The analog signal chain consists of a
pair of high- and low-pass filters (Mini-Circuits ZFHP-1R2+ and
SLP-150+) that together band-limit the signal to 1.2--140~MHz, and the
filters are followed by a Mini-Circuits ZFL-500+ amplifier.  In
contrast to the two-element pathfinder, the low-pass cutoff is
increased from 81~MHz to 140~MHz to capture downlink signals at
137--138~MHz from the ORBCOMM satellite constellation.  For the final
\albatros\ array, the ORBCOMM signals provide a convenient means for
synchronization across the antenna stations, serving as a backup to
the GPS timing discussed below.  Preliminary lab tests suggest that on
time scales of $\sim30$~s, relative timing between different antenna
stations can be measured to an accuracy of better than a few tenths of
a nanosecond using a single satellite.  The corresponding phase error
at 10~MHz is $\lesssim1^\circ$.  With publicly available orbits and
multiple satellites typically within the field of view, we expect that
ORBCOMM baseband data saved at the same time as the astronomical data
can provide offline synchronization of the \albatros\ stations.
Because the ORBCOMM and science data are taken at the same time by the
same system, improvements to the ORBCOMM calibration can be
immediately carried over into improving the science data calibration,
even if said improvements take years to realize.

As with the two-element pathfinder, a SNAP board digitizes each of the
two RF signals at 250~Msamp/s. In the autonomous station
configuration, the SNAP board ADCs are locked to a 10-MHz reference
produced by a Trimble Thunderbolt~E GPS-disciplined clock module.  The
SNAP board FPGA computes two data products: 1) channelized baseband
data for each polarization over tunable frequency windows within the
0--125~MHz operating range, with the options of 1-, 2-, or 4-bit
compression, and 2) auto- and cross-spectra from the two polarizations
over the full 0--125~MHz span, accumulated over few-second intervals.
The reduction in bit depth happens only \textit{after} the SNAP board
has channelized the baseband, and so cross-channel leakage (due to
\textit{e.g.} RFI and slopes in RF power as seen by the ADCs) is
unaffected by the low bit depth.  
An RPi~3B+ controls the SNAP board and and receives the auto- and
cross-spectra via GPIO connections, and the spectra are saved to an
on-board SD card.  The baseband data are passed from the SNAP board to
the RPi via ethernet and written to external hard drives.  The
introduction of gigabit ethernet with the RPi~3B+ model has enabled
the high data throughput associated with writing baseband.  As a
benchmark, 1-bit baseband recording of two polarizations over 10~MHz
of bandwidth yields an approximate data rate of 5~MB/s, or 0.4~TB/day.

@@

@@

@@

@@

@@

We gratefully acknowledge the National Research Foundation (grant
number 110929), the South African National Antarctic Programme, the
Natural Sciences and Engineering Research Council of Canada (grant
number RGPNS-2019-534549), and the Polar Continental Shelf Program for
providing funding and logistical support for our research program.
This research was undertaken, in part, thanks to funding from the
Canada 150 Program.  The financial assistance of the South
African SKA Project (SKA SA) towards this research is hereby
acknowledged (\url{www.ska.ac.za}).  We additionally extend our
sincere gratitude to the staff at SKA SA for hosting our lab tests.
We thank the South African National Space Agency for their technical
support, the crew of Ultimate Heli for safely delivering us and our
cargo, and all of the Marion takeover and winter team members who
provided invaluable advice and field help.  The authors also wish to
thank members of the LWA and RAPID collaborations for useful
discussions, and Raul Monsalve for helping make our first MARS
campaign a success.
